---
title: "Compare  results model 3 and model 4"
author: "Willem Vervoort, Michaela Dolk & Floris van Ogtrop"
date: "`r Sys.Date()`"
output: 
    pdf_document:
      fig_width: 7
      fig_height: 6
      fig_caption: true
---
```{r packages, warning=F, message=F}
require(tidyverse)
require(lubridate)
require(pander)
panderOptions('table.split.cells', 12)
panderOptions('table.alignment.default', 'center')
panderOptions('table.alignment.rownames', 'right')
```
```{r setup}
# root dir
knitr::opts_knit$set(root.dir = "C:/Users/rver4657/ownCloud/Virtual Experiments/VirtExp")
```

This rmarkdown document and the resulting pdf are stored on  [github](https://github.com/WillemVervoort/VirtExp). All directories (apart from the root working directory) refer to the directories in this repository

## Introduction
This document is related to the manuscript "Disentangling climate change trends in Australian streamflow" (vervoort et al.). This document answers the reviewers question whether there is a systematci difference in the AIC values for "model 3" and "model 4", see document number 3 in this series.

## Read in the data  
The data are the AIC values from Table 5 and Table 6 in the manuscript.
```{r readdata}
Table5 <- read_csv("../ProjectData/Table5.csv")
Table6 <- read_csv("../ProjectData/Table6.csv")
```

Extract the AIC values and combine into a new table

```{r AICtable}
AICTable <- tibble(Catchment = Table5$Station,
                   AICModel3 = Table5$AIC,
                   AICModel4 = Table6$AIC,
                   `Difference3-4` = Table5$AIC - Table6$AIC)

pander(AICTable)

mean(AICTable$`Difference3-4`)
median(AICTable$`Difference3-4`)
```

## plot on 1:1

Plot the AIC values against each other and insert a 1:1 line

```{r AICplot}
AICTable %>%
  ggplot(aes(AICModel3,AICModel4,colour=Catchment)) +
  geom_point(size=3) + geom_abline(linetype = 2) + theme_bw()


```

