---
title: "Back transformation theory"
author: "Willem Vervoort, Michaela Dolk & Floris van Ogtrop"
date: "`r Sys.Date()`"
output: 
    pdf_document:
      fig_width: 7
      fig_height: 6
      fig_caption: true
---
```{r setup, warning=F, message=F}
# root dir
knitr::opts_knit$set(root.dir = 
                       "D:/Cloudstor/Virtual Experiments/VirtExp")
                       #"C:/Users/rver4657/ownCloud/Virtual Experiments/VirtExp")
knitr::opts_chunk$set(echo = TRUE)
# LOAD REQUIRED PACKAGES # #####
library(pander)
library(tidyr)
library(zoo)
library(mgcv)
library(ggplot2)
```
This rmarkdown document and the resulting pdf are stored on  [github](https://github.com/WillemVervoort/VirtExp). All directories (apart from the root working directory) refer to the directories in this repository. 

# Introduction
This document is related to the manuscript "Disentangling climate change trends in Australian streamflow" (vervoort et al.), submitted.

# Back transformation
This document is related to the *[3.GAMmodelTest.pdf](./3.GAMmodelTests.pdf)* document and is to explain the back transformation used in the `gamm()` and `gls()` regressions, which are essentially the following type of regression equation  
$log(y) = \beta_1*trend$  
How do we interpret $\beta_1$?  
Assume a 1 unit change in trend in time, this results in:  
$log(y_{t=1}) - log(y_{t=0}) = \beta_1$ or  
$log(\frac{y_{t=1}}{y_{t=0}}) = \beta_1$  
raising both sides to a power results in:  
$\frac{y_{t=1}}{y_{t=0}} = exp(\beta_1)$  

[According to this link](https://data.library.virginia.edu/interpreting-log-transformations-in-a-linear-model/), we need to subtract 1 (for log(y + 1) = 0, presumably) to get the ratio

In other words, $exp(\beta_1) - 1$ is equal to the ratio of the y values, or can be interpreted as a fractional change in y. The reason for substracting one is that we are talking about a 1 unit increase in the trend results in an $exp(\beta_1)$ change in y, so to get to fractional change we need to subtract 1

The addition of a simple scalar  to y (1 in this case) makes no difference on the trend interpretation. This simply means we are now looking at a change in y+1, which is essentially the same as the change in y as the addition is a scalar.


```{r small_demo}
load("data/ClimCh_project_MD.Rdata")

test <- flow_rain_maxT_weekly[flow_rain_maxT_weekly$Station=="DOMB",]
# add a vector of trend values
test$trend <- 1:nrow(test)


# plot distributions
par(mfrow=c(2,2))
hist(test$Flow, main="untransformed")
hist(log10(test$Flow +1), main ="log10 trans + 1")
hist(log1p(test$Flow), main = "log(flow+1)")

# log10 seems best
# modelling
plot(log10(Flow + 1)~trend, data = test)
mod1 <- lm(log10(Flow + 1)~trend, data = test)
summary(mod1)
plot(residuals(mod1))
# interpret the slope
slope <- exp(coef(mod1)[2])-1
slope
```

This slope is the fractional change in the flow per week. So if we multiply by 100 we get the "average weekly % change" in the flow.
This implies the same % change on an annual scale and can be straight away interpreted as the amplification

This suggest a 0.005% decrease in flow
