---
title: "Extract Adj r-squared GAMM model 5"
author: "Willem Vervoort, Michaela Dolk & Floris van Ogtrop"
date: "`r Sys.Date()`"
output: 
    pdf_document:
      fig_width: 7
      fig_height: 6
      fig_caption: true
---

```{r setup, warning=F, message=F}
# root dir
knitr::opts_knit$set(root.dir = 
                       "d:/cloudstor/Virtual Experiments/VirtExp")
knitr::opts_chunk$set(echo = TRUE)
# LOAD REQUIRED PACKAGES # #####
library(pander)
library(tidyverse)
library(xts)
library(zoo)
library(mgcv)
library(Kendall)
library(doParallel)
library(foreach)

storedir <- "d:/cloudstor/virtual experiments"

```


This rmarkdown document and the resulting pdf are stored on  [github](https://github.com/WillemVervoort/VirtExp). All directories (apart from the root working directory) refer to the directories in this repository

# Introduction
This document is related to the manuscript "Disentangling climate change trends in Australian streamflow" (vervoort et al.), submitted to Journal of Hydrology.
This section only extracts the performance results (Ajusted r_squared) from the GAMM results of model 4 and 5 from part 3 of the series, [3.GAMmodelTests.pdf](./3.GAMmodelTests.pdf)

# Read in the results

```{r}

# currently not saved
# # Station rainfall model 4
# load(paste(storedir,
#                 "projectdata/StoreFwRE_TrendAnalysis.RData",
#                 sep = "/"))


#Gridded raifnall model 4
load(paste(storedir,
                "projectdata/StoreFwGRE_TrendAnalysis.RData",
                sep="/"))

# Station rainfall model 5
load(paste(storedir,
                "projectdata/StoreFwRE_Analysis.RData",
                sep="/"))

# Gridded Rainfall Model 5
load(paste(storedir,
                "projectdata/StoreFwGRE2_TrendAnalysis.RData",
                sep="/"))

```


# extract the adjusted r-squared

```{r}

# Gridded rainfall model 4
Model4Grid_AdjR2 <- list()

for (i in 1:13) {
  Model4Grid_AdjR2[[i]] <- summary(Store_FwGRE[[i]]$model$gam)$r.sq
  
}

Model4Grid_AdjR2 <- do.call(rbind,Model4Grid_AdjR2)

# Station rainfall model 5
Model5Station_AdjR2 <- list()

for (i in 1:13) {
  Model5Station_AdjR2[[i]] <- summary(Store_FwRE2[[i]]$model$gam)$r.sq
  
}

Model5Station_AdjR2 <- do.call(rbind,Model5Station_AdjR2)



# Gridded rainfall model 5
Model5Grid_AdjR2 <- list()

for (i in 1:13) {
  Model5Grid_AdjR2[[i]] <- summary(Store_FwGRE2[[i]]$model$gam)$r.sq
  
}

Model5Grid_AdjR2 <- do.call(rbind,Model5Grid_AdjR2)

Results <- tibble(Station= do.call(rbind, lapply(1:length(Store_FwGRE), 
                          function(i) rbind(Store_FwGRE[[i]][[2]][[1]])))$Station,
                  Model4_Grid = Model4Grid_AdjR2, Model5_Station = Model5Station_AdjR2,
                  Model5_Grid = Model5Grid_AdjR2)
pander(Results, caption="Adjusted r-squared for models 4 and 5 from the GAMM analysis")

```



