---
title: "4. FDC comparison"
author: "Willem Vervoort, Michaela Dolk & Floris van Ogtrop"
date: "`r Sys.Date()`"
output: 
    pdf_document:
      fig_caption: yes
      fig_height: 8
      keep_tex: yes
---
```{r setup, warning=F, message=F}
# root dir
knitr::opts_knit$set(root.dir = "C:/Users/rver4657/ownCloud/Virtual Experiments/VirtExp")
knitr::opts_chunk$set(echo = TRUE)
# LOAD REQUIRED PACKAGES # #####
library(pander)
library(tidyr)
library(xts)
library(zoo)
library(ggplot2)
library(reshape2)
```

This rmarkdown document and the resulting pdf are stored on  [github](https://github.com/WillemVervoort/VirtExp). All directories (apart from the root working directory) refer to the directories in this repository

# Introduction
This document is related to the manuscript "Disentangling climate change trends in Australian streamflow" (vervoort et al.), submitted to Journal of Hydrology. This is the fourth document as part of the response to reviewers of earlier versions of the manuscript outlining in detail the generation of the flow duration curves. This is **Figure 6 and Figure 7** in the manuscript.
The key idea in this analysis is that by comparing the flow duration curves by decade, we might be able to pick up where exactly the rainfall and the flow distribution have changed. An alternative analysis of the rainfall using double mass curves is also presented.

# Load data and find decades

```{r loaddata}
load("data/DailyDataIncludingGridded.Rdata")
load("data/ClimCh_project_MD.Rdata")
# correct the column name of maxT in GridRainAllDataout
colnames(GridRainAllDataout)[5] <- "MaxT"
# define decade start and end
decade_start <- c(as.Date("1/1/1970", format="%d/%m/%Y"), 
                  as.Date("1/1/1980", format="%d/%m/%Y"), 
                  as.Date("1/1/1990", format="%d/%m/%Y"), 
                  as.Date("1/1/2000", format="%d/%m/%Y"))
decade_end <- c(as.Date("31/12/1979", format="%d/%m/%Y"), 
                as.Date("31/12/1989", format="%d/%m/%Y"), 
                as.Date("31/12/1999", format="%d/%m/%Y"), 
                as.Date("31/12/2009", format="%d/%m/%Y"))

```

The next step is to assign the decades to the daily data as the weekly data already have associated decades. In this document we will plot both the weekly and the daily flow duration curves (FDC) to check if aggregating to weekly makes a difference.

```{r findDecades}
flow_decade <- vector("character", length=nrow(flow_zoo))
decades <- c("70-80","80-90", "90-00", "00-10")
for(i in 1:4) {
  flow_decade[time(flow_zoo) >= decade_start[i] & 
                    time(flow_zoo) <= decade_end[i]] <- decades[i]
}
```

# Flow duration curves for the rainfall and the flow data
Flow durations curves are essentially a frequency curve and therefor we can apply the same analysis to the rainfall and the flow data. The best would be to use all the data and rank this to generate the FDC, but this is not possible as there are different numbers of NA values. So here we use the function `quantile()` to generate the distributions. For the daily data we generate a series of 1000 samples, while for the weekly data, we generate 100 samples.

## flow data

```{r FDC_flow, fig.cap="Difference in flow duration curves based on daily data split by decade"}
# combine flow_zoo with flow_decade in a data.frame
flow_df <- data.frame(decade=flow_decade,
                         as.data.frame(flow_zoo))
# now calculate the flow duration curve for each catchment by decade
# function to generate FDC, use quantile decades have different data lenghts (leap years)
FDC_gen <- function(DATA) {
   FDC <- data.frame(probs = seq(0,1,length=1000)*100,
                     flow = quantile(DATA,probs=seq(0,1,length=1000),
                                     na.rm=T))
    return(FDC)
}

# tapply this across the columns and produce a list
# do this for each decade
# an empty list for the decades
FDCs <- list()
for (i in 1:4) {
  FDCs[[i]] <- apply(subset(flow_df,
                as.character(flow_df$decade)==
                  as.character(decades[i]))[,2:14],
                      2,FDC_gen)
}

colour_list <- c("red", "purple", "blue", "green")
# DIFFERENCE BETWEEN WEEKLY FDCs DIVIDED BY 1970-1979 FDCs -> not log, ylim=c(-1,1) to prevent issues plotting when denominator 0 (gives -Inf) hence not all values shown

periods <- c("((1970-1979) - (1980-1989))/(1970-1979)",
              "((1970-1979) - (1990-1999))/(1970-1979)",             
              "((1970-1979) - (2000-2009))/(1970-1979)")

plot.list <- vector("list", length=13)
temp <- list()
for (j in 1:length(periods)) {
  for (i in seq_along(Stations[,1])) {
    temp[[i]] <- (FDCs[[1]][[i]]$flow - FDCs[[j+1]][[i]]$flow)/FDCs[[1]][[i]]$flow
  }
  fdc <- do.call(cbind,temp)
  colnames(fdc) <- Stations[,1]
  plot.list[[j]] <- melt(data.frame(fdc = fdc, 
                                period = rep(periods[j],nrow(fdc))),
                                id.vars="period")
}
# bring back together to a dataframe
plot.df <- do.call(rbind,plot.list)
# add the probabilities
plot.df <- cbind(plot.df,probs=rep(FDCs[[1]][[1]]$probs,13*3))
colnames(plot.df)[2] <- "station"
levels(plot.df$station) <- Stations[,1]

# make a plot
p <- ggplot(plot.df, aes(x = probs, y = value)) +
  geom_line(aes(linetype=period, colour=period),size=1.2) + 
  facet_wrap(~ station,ncol=5) + ylim(c(-2,2)) +
  theme(legend.position="bottom") +
  guides(col = guide_legend(nrow = 3))
p <- p + ggtitle("Streamflow") +
  theme(plot.title = element_text(lineheight=.8, face="bold")) +
  xlab("Probability") +
  theme(axis.title.x = element_text(face="bold",  size=16),
        axis.text.x  = element_text(size=12)) +
  ylab("Scaled Difference") +
  theme(axis.title.y = element_text(face="bold",  size=16),
        axis.text.y  = element_text(size=12)) +
 scale_colour_manual(values=c("black", "gray33", "gray66")) +
  theme(legend.text = element_text( size = 12))+
  theme(legend.title = element_text(size=14, face="bold")) +
  theme(strip.text.x = element_text(size=16))
print(p)

# # publication quality
# tiff("../manuscript/Figure6_StreamflowFDCDifference.tif",
#      width=16*480, height=12*480,
#      res=600, compression="lzw")
# print(p)
# dev.off()
```

## daily rainfall data
```{r FDC_rain, fig.cap="Difference in rainfal duration curves based on daily data split by decade"}
# combine flow_zoo with flow_decade in a data.frame
rain_df <- data.frame(decade=flow_decade,
                         as.data.frame(rain_zoo))
# now calculate the rainfall duration curve for each catchment by decade
# tapply this across the columns and produce a list
# do this for each decade
# an empty list for the decades
FDCs <- list()
for (i in 1:4) {
  FDCs[[i]] <- apply(subset(rain_df,
      as.character(rain_df$decade)==as.character(decades[i]))[,2:14],
                      2,FDC_gen)
}

# DIFFERENCE BETWEEN WEEKLY FDCs DIVIDED BY 1970-1979 FDCs -> not log, ylim=c(-1,1) to prevent issues plotting when denominator 0 (gives -Inf) hence not all values shown
plot.list <- vector("list", length=13)
temp <- list()
for (j in 1:length(periods)) {
  for (i in seq_along(Stations[,1])) {
    temp[[i]] <- (FDCs[[1]][[i]]$flow - FDCs[[j+1]][[i]]$flow)/FDCs[[1]][[i]]$flow
  }
  fdc <- do.call(cbind,temp)
  colnames(fdc) <- Stations[,1]
  plot.list[[j]] <- melt(data.frame(fdc = fdc, 
                                period = rep(periods[j],nrow(fdc))),
                                id.vars="period")
}
# bring back together to a dataframe
plot_df_r <- do.call(rbind,plot.list)
# add the probabilities
plot_df_r <- cbind(plot_df_r,probs=rep(FDCs[[1]][[1]]$probs,13*3))
colnames(plot_df_r)[2] <- "station"
levels(plot_df_r$station) <- Stations[,1]

# make a plot
p <- ggplot(plot_df_r, aes(x = probs, y = value)) +
  geom_line(aes(linetype=period, colour=period),size=1.2) + 
  facet_wrap(~ station,ncol=5) + ylim(c(-2,2)) +
  theme(legend.position="bottom") +
  guides(col = guide_legend(nrow = 3))
p <- p + ggtitle("Rainfall") +
  theme(plot.title = element_text(lineheight=.8, face="bold"))+
  xlab("Probability") +
  theme(axis.title.x = element_text(face="bold",  size=16),
        axis.text.x  = element_text(size=12)) +
  ylab("Scaled Difference") +
  theme(axis.title.y = element_text(face="bold",  size=16),
        axis.text.y  = element_text(size=12)) +
 scale_colour_manual(values=c("black", "gray33", "gray66")) +
  theme(legend.text = element_text( size = 12))+
  theme(legend.title = element_text(size=14, face="bold")) +
  theme(strip.text.x = element_text(size=16))
print(p)

# publication quality
# tiff("../manuscript/Figure7_RainfallFDCDifference.tif",
#      width=16*480, height=12*480,
#      res=600, compression="lzw")
# print(p)
# dev.off()
```

## daily gridded rainfall
```{r FDC_gridRain, fig.cap="Difference in gridded rainfal duration curves based on daily gridded data split by decade"}
# gridded rainfall
grrain_df <- data.frame(decade=rep(flow_decade,13),
                         as.data.frame(GridRainAllDataout[,1:2]))
# now calculate the rainfall duration curve for each catchment by decade
# tapply this across the columns and produce a list
# do this for each decade
# an empty list for the decades
FDCs <- list()
for (i in 1:4) {
  temp <- subset(grrain_df,
                        as.character(grrain_df$decade)==  
                        as.character(decades[i]))
  FDCs[[i]] <- tapply(temp[,"gridRain"],
                      list(Station=temp$Station),FDC_gen)
}

# plotting
plot.list <- vector("list", length=13)
temp <- list()
for (j in 1:length(periods)) {
  for (i in seq_along(Stations[,1])) {
    temp[[i]] <- (FDCs[[1]][[i]]$flow - FDCs[[j+1]][[i]]$flow)/FDCs[[1]][[i]]$flow
  }
  fdc <- do.call(cbind,temp)
  colnames(fdc) <- Stations[,1]
  plot.list[[j]] <- melt(data.frame(fdc = fdc, 
                                period = rep(periods[j],nrow(fdc))),
                                id.vars="period")
}
# bring back together to a dataframe
plot_df_r <- do.call(rbind,plot.list)
# add the probabilities
plot_df_r <- cbind(plot_df_r,probs=rep(FDCs[[1]][[1]]$probs,13*3))
colnames(plot_df_r)[2] <- "station"
levels(plot_df_r$station) <- Stations[,1]

# make a plot
p <- ggplot(plot_df_r, aes(x = probs, y = value)) +
  geom_line(aes(linetype=period, colour=period),size=1.2) + 
  facet_wrap(~ station,ncol=5) + ylim(c(-2,2)) +
  theme(legend.position="bottom") +
  guides(col = guide_legend(nrow = 3))
p <- p + ggtitle("gridded Rainfall") +
  theme(plot.title = element_text(lineheight=.8, face="bold"))+
  xlab("Probability") +
  theme(axis.title.x = element_text(face="bold",  size=16),
        axis.text.x  = element_text(size=12)) +
  ylab("Scaled Difference") +
  theme(axis.title.y = element_text(face="bold",  size=16),
        axis.text.y  = element_text(size=12)) +
 scale_colour_manual(values=c("black", "gray33", "gray66")) +
  theme(legend.text = element_text( size = 12))+
  theme(legend.title = element_text(size=14, face="bold")) +
  theme(strip.text.x = element_text(size=16))
print(p)
```

