---
title: "Monthly data GAM and data preparation"
author: "Willem Vervoort, Michaela Dolk & Floris van Ogtrop"
date: "`r Sys.Date()`"
output: 
    pdf_document:
      fig_width: 7
      fig_height: 6
      fig_caption: true
---
```{r setup, warning=F, message=F}
# root dir
knitr::opts_knit$set(root.dir = "C:/Users/rver4657/ownCloud/Virtual Experiments/VirtExp")
knitr::opts_chunk$set(echo = TRUE)
# LOAD REQUIRED PACKAGES # #####
library(pander)
library(tidyverse)
library(zoo)
library(ggplot2)
library(hydromad)
library(Kendall)
library(mgcv)
library(doParallel)
library(foreach)
library(xts)
```

This rmarkdown document and the resulting pdf are stored on  [github](https://github.com/WillemVervoort/VirtExp). All directories (apart from the root working directory) refer to the directories in this repository.

# Introduction
This document is related to the manuscript "Disentangling climate change trends in Australian streamflow" (vervoort et al.), submitted to Journal of Hydrology. This is the eight^th^ part of the series that reruns the Mann Kendall and GAMM analysis on the monthly station data and creates monthly output for running the numerical modelling with monthly data.

So the steps that we need to do are:

- Aggregate the station data to monthly;
- run the Mann Kendall and GAMM analysis;
- write the monthly data as a datafile to be read in on the HPC modelling. 

However, the monthly Mann-Kendall analysis has already been run in the 2^nd^ part of the series (2.MannKendallTest.pdf), so we only need to run the GAMM analysis

# Aggregate station data to monthly


```{r loaddata}
load("data/ClimCh_project_MD.Rdata")


flow_zoo_m <- aggregate(flow_zoo,as.yearmon,sum,na.rm=T)
rain_zoo_m <- aggregate(rain_zoo,as.yearmon,sum,na.rm=T)
maxT_zoo_m <- aggregate(maxT_zoo,as.yearmon,sum,na.rm=T)

flow_rain_maxT_monthly <- melt(as.data.frame(rain_zoo_m))
colnames(flow_rain_maxT_monthly) <- c("Station","Rain")
flow_rain_maxT_monthly$Station <-
  substr(flow_rain_maxT_monthly$Station,1,4)

flow_rain_maxT_monthly$Flow <- melt(as.data.frame(flow_zoo_m))[,2]
flow_rain_maxT_monthly$MaxT <- melt(as.data.frame(maxT_zoo_m))[,2]

rm(list=c("flow_rain_maxT_weekly","GridRainAllDataout",
     "weekGridRainAllDataout",
     "flow_zoo","maxT_zoo", "rain_zoo"))

```

# Run the GAMM analysis on the monthly data

## Model 1 Only flow and trend
The first 2 models are actually not generalised additive mixed models (GAM) as the models only analyse a linear trend. To match the GAM analysis, we used generalised least squares (`gls()`) in R. This still allows correlated errors to be analysed

```{r model1, fig.cap="Residuals of linear mixed model analysis for trend in monthly flow only"}
# run the gls model on flowtrend only
cl <- makeCluster(4) # create a cluster with 4 cores
registerDoParallel(cl) # register the cluster
# use a foreach loop to calibrate
Store2 <- foreach(i = 1:length(Stations[,1]),
                 .packages="mgcv") %dopar% {
  gamm.data <- subset(flow_rain_maxT_monthly,
                      flow_rain_maxT_monthly$Station == Stations[i,1])
  gamm.data$trend <- 1:nrow(gamm.data)
  gam_TrendOnly <- gls(log(Flow +1)~trend, correlation= corCAR1(),
       data=gamm.data)
  out <- list(model = gam_TrendOnly,
              results = data.frame(Station=Stations[i,1],
              t(summary(gam_TrendOnly)$tTable[2,c(1,4)]),
                   AIC=summary(gam_TrendOnly)$AIC))
  out
}
stopCluster(cl)

par(mfrow=c(5,3),mar=c(2,2,2,2))
for (i in seq_along(Stations[,1])) {
  res <- residuals(Store2[[i]]$model)  
  plot(res, main=Stations[i,1], cex.main=0.7,
       ylab="normalised residuals",xlab="")
  n <- length(res)
  abline(lsfit(1:n, res), col="red")
}

# store results
storedir <- "c:/users/rver4657/owncloud/virtual experiments"
save(Store2,file=paste(storedir,
                "projectdata/Store2_MonthlyTrendOnly.RData",
                sep="/"))
output <- do.call(rbind, lapply(1:length(Store2), function(i) rbind(Store2[[i]][[2]])))
pander(output, caption="Mixed model results for analysis of trend in monthly flow only")
rm(Store2)
```

## Model 2 trend in rain  
Similar to the flow data, this analysis uses `gls()` to run the linear mixed model to test for a trend in the data and compare to the Mann-Kendall results

```{r model2a, fig.cap="Residuals of linear mixed model analysis for trend in monthly station rainfall data"}
# create an empty list
# and an empty dataframe to store results
# run the gls model on flowtrend only
cl <- makeCluster(4) # create a cluster with 4 cores
registerDoParallel(cl) # register the cluster
# use a foreach loop to calibrate
Store_Rain <- foreach(i = 1:length(Stations[,1]),
                 .packages="mgcv") %dopar% {
  gamm.data <- subset(flow_rain_maxT_monthly,
                      flow_rain_maxT_monthly$Station == Stations[i,1])
  gamm.data$trend <- 1:nrow(gamm.data)
  gam_TrendR <- gls(log(Rain + 1)~trend, correlation= corCAR1(),
       data=na.omit(gamm.data))
  out <- list(model = gam_TrendR,
              results = data.frame(Station=Stations[i,1],
                     t(summary(gam_TrendR)$tTable[2,c(1,4)]),
                                   AIC=summary(gam_TrendR)$AIC))
  out
}
stopCluster(cl)


par(mfrow=c(5,3),mar=c(2,2,2,2))
for (i in seq_along(Stations[,1])) {
  res <- residuals(Store_Rain[[i]]$model)  
  plot(res, main=Stations[i,1], cex.main=0.5,
       ylab="normalised residuals",xlab="")
  n <- length(res)
  abline(lsfit(1:n, res), col="red")
}

# store results
save(Store_Rain,file=paste(storedir,
                  "projectdata/StoreRain_MonthlyTrend.RData",
                    sep="/"))
output <- do.call(rbind, lapply(1:length(Store_Rain), 
                     function(i) rbind(Store_Rain[[i]][[2]])))
pander(output, caption="Mixed model results for analysis of trend in Monthly Station Rainfall")
rm(Store_Rain)
```

## Model 3 GAMM with rainfall  
This model analyses flow as a function of rainfall only. This is therefore an analysis of the rainfall runoff coefficient, taking into account a possible time trend in the data. If the trend in this analysis is significant, then this is a measure of how the rainfall runoff coefficient has changed over time.  

```{r model3a, fig.cap="Residuals of GAMM analysis for trend in monthly flow data taking into consideration station rainfall data"}
# Gamm model with flow and rain
cl <- makeCluster(4) # create a cluster with 4 cores
registerDoParallel(cl) # register the cluster
# use a foreach loop to calibrate
Store_FwR <- foreach(i = 1:length(Stations[,1]),
                 .packages="mgcv") %dopar% {
  gamm.data <- subset(flow_rain_maxT_monthly,
                      flow_rain_maxT_monthly$Station == Stations[i,1])
  gamm.data$trend <- 1:nrow(gamm.data)
  gam_TrendFlow_withR <- gamm(log(Flow +1)~s(Rain) + trend, 
                                  correlation= corCAR1(), data=gamm.data)
  out <- list(model = gam_TrendFlow_withR,
              results = data.frame(Station=Stations[i,1],
                            t(summary(gam_TrendFlow_withR$lme)$tTable[2,c(1,5)]),
                                   AIC=summary(gam_TrendFlow_withR$lme)$AIC))
  out
   }
stopCluster(cl)

par(mfrow=c(5,3), mar=c(2,2,2,2))
for (i in seq_along(Stations[,1])) {
  res <- residuals(Store_FwR[[i]]$model$lme)  
  plot(res, main=Stations[i,1], cex.main=0.5,
       ylab="normalised residuals",xlab="")
  n <- length(res)
  abline(lsfit(1:n, res), col="red")
}

# store results
save(Store_FwR,
     file=paste(storedir,
              "projectdata/StoreFwR_monthlyTrend.RData",
              sep="/"))
output <- do.call(rbind, lapply(1:length(Store_FwR), 
                                function(i) rbind(Store_FwR[[i]][[2]])))
pander(output, caption="Mixed model results for analysis of trend in monthly flow data taking into account Rainfall")
rm(Store_FwR)
```

## Model 4. GAMM with rain & s(rain,MaxT) and trend
This model analyses flow as a function of rainfall and the interaction between rainfall and maximum temperature, which is conceptualised as the actual evapotranspiration. This is therefore an analysis of the rainfall runoff coefficient, taking into account the changes in evapotranspiration and a possible time trend in the data. If the trend in this analysis is significant, then this is a measure of how the rainfall runoff coefficient has changed over time.  


```{r model4a, fig.cap="Residuals of GAMM analysis for trend in monthly flow data removing station rainfall and evapotranspiration effects"}
# run the gamm model on rain, maxT and flow
cl <- makeCluster(4) # create a cluster with 4 cores
registerDoParallel(cl) # register the cluster
# use a foreach loop to calibrate
Store_FwRE <- foreach(i = 1:length(Stations[,1]),
                 .packages="mgcv") %dopar% {
  gamm.data <- subset(flow_rain_maxT_monthly,
                      flow_rain_maxT_monthly$Station == Stations[i,1])
  gamm.data$trend <- 1:nrow(gamm.data)
  gam_TrendFlow_withRandE <- gamm(log(Flow +1)~s(Rain) + s(Rain, MaxT) + 
                                       trend, correlation= corCAR1(),
                                       data=gamm.data, control=list(niterEM=0))
  out <- list(model = gam_TrendFlow_withRandE,
        results = data.frame(Station=Stations[i,1],
                      t(summary(gam_TrendFlow_withRandE$lme)$tTable[2,c(1,5)]),
                            AIC=summary(gam_TrendFlow_withRandE$lme)$AIC))
  out
 }
stopCluster(cl)

par(mfrow=c(5,3), mar=c(2,2,2,2))
for (i in seq_along(Stations[,1])) {
  res <- residuals(Store_FwRE[[i]]$model$lme)  
  plot(res, main=Stations[i,1], cex.main=0.5,
       ylab="normalised residuals",xlab="")
  n <- length(res)
  abline(lsfit(1:n, res), col="red")
}

# store results
save(Store_FwRE,
     file=paste(storedir,
                "projectdata/StoreFwRE_MonthlyTrend.RData",
                sep = "/"))
output <- do.call(rbind, lapply(1:length(Store_FwRE), 
                                function(i) rbind(Store_FwRE[[i]][[2]])))
pander(output, caption="Mixed model results for the analysis of trend in monthly flow data taking into account Rainfall and Evapotranspiratiion")
rm(Store_FwRE)
```

## Model 5, same as model 4, but no trend and Mann Kendall on the residuals
This last model is to check the trend with GAMM analysis with the analysis using Mann-Kendall. So rather than incorporating a trend in the model, we analyse the residuals using Mann-Kendall for a trend. In this case we drop the plotting of the residuals.

```{r model5a}
# run the gamm model on rain, maxT and flow
cl <- makeCluster(4) # create a cluster with 4 cores
registerDoParallel(cl) # register the cluster
# use a foreach loop to calibrate
Store_FwRE2 <- foreach(i = 1:length(Stations[,1]),
                 .packages="mgcv") %dopar% {
  gamm.data <- subset(flow_rain_maxT_monthly,
                      flow_rain_maxT_monthly$Station == Stations[i,1])
  gam_Flow_withRandE <- gamm(log(Flow +1)~s(Rain) + s(Rain, MaxT) , 
                                  correlation= corCAR1(), data=gamm.data,
                                  control=list(niterEM=5))
  out <- list(model = gam_Flow_withRandE,
        results = data.frame(Station=Stations[i,1],
                            AIC=summary(gam_Flow_withRandE$lme)$AIC))
  out
}

stopCluster(cl)

# store results
save(Store_FwRE2,
     file=paste(storedir,
                "projectdata/StoreFwRE_Monthly.RData",
                sep="/"))
output <- do.call(rbind, lapply(1:length(Store_FwRE2), 
                                function(i) rbind(Store_FwRE2[[i]][[2]])))
pander(output, caption="Mixed model results for the analysis of monthly flow data taking into account Rainfall and Evapotranspiration")
```

Now do the Mann-Kendall analysis on the residuals  

```{r model5aMK, fig.cap="Mann Kendall analysis of the residuals of the monthly streamflow after GAM model with rainfall and a Evapotranspiration"}
# do mann kendall on the residuals
resid_list <- vector("list", length=length(Stations[,1]))
for (i in seq_along(Stations[,1])) {
  resid_list[[i]] <- zoo(residuals(Store_FwRE2[[i]]$model$lme,
                 type="normalized"),
          order.by=time(flow_zoo_m))
}
resid_df <- do.call(merge.zoo,resid_list)
names(resid_df) <- Stations[,1]
# Bootstrap
set.seed(10)
# now run a loop over the number of years (create 41 different sets)
# do Mann Kendall test on each resonstituted series
# ---------------------------
#  -------------------------
resid_temp <- as.data.frame(resid_df)
resid_temp$years <- format(time(resid_df),"%Y")
split_resid <- split(resid_temp[,1:13],resid_temp$years)


cl <- makeCluster(4) # create a cluster with 4 cores
registerDoParallel(cl) # register the cluster
# use a foreach loop to calibrate
MK_list <- foreach(i = 1:500,
                 .packages=c("Kendall", "xts")) %dopar% {
  # reorganise the list elements
  series <- sample(1:nyears(resid_df),nyears(resid_df))
  for (j in 1:length(series)) {
    if (j==1) { 
      new_df <- as.data.frame(split_resid[[series[j]]])
    } else {
      new_df <- rbind(new_df,as.data.frame(split_resid[[series[j]]]))
    }
  }
  # run mann kendall on the columns and store the results
  mk_r <- apply(new_df,2,MannKendall)
  
  out <- do.call(cbind,mk_r)
  out
 }
stopCluster(cl)


MK_df <- do.call(rbind,MK_list)

pvalues <- subset(MK_df, rownames(MK_df)=="sl")
tau <- subset(MK_df, rownames(MK_df)=="tau")

sig_set <- list()

for (i in 1:ncol(pvalues)) {
  set <- data.frame(pvalue=as.numeric(pvalues[,i]),
                    tau=as.numeric(tau[,i]),
                    catch=rep(colnames(MK_df)[i],nrow(tau)))
  sig_set[[i]] <- set[set$pvalue < 0.5,]
}

sig_set_a <- do.call(rbind,sig_set)
sig_set_a$type <- rep("bootstrap",nrow(sig_set_a))


MK_resid <- do.call(rbind,lapply(resid_list,MannKendall))


real_df <- data.frame(pvalue = as.numeric(MK_resid[,2]), 
                      tau = as.numeric(MK_resid[,1]),
                      catch=Stations[,1],
                      type=rep("real",nrow(MK_resid)))
# A histogram of taus

hp <- ggplot(sig_set_a, aes(x=tau)) + geom_histogram(binwidth=0.03,colour="white")
# Histogram of significant tau's, divided by catch
# With panels that have the same scaling, but different range 
# (and therefore different physical sizes)
hp <- hp + facet_wrap(~ catch,ncol=5)
# add a red point for the real slope from the data
p_value <- ifelse(real_df$pvalue<0.05,"< 0.05",">= 0.05")
hp <- hp + geom_point(data=real_df,aes(x=tau, y=0,colour=p_value),
                shape=16,size=5) + 
  facet_wrap(~ catch,ncol=5)+ ggtitle("Residuals Streamflow after GAM") #+
hp <- hp + scale_colour_grey(start = 0, end = 0.6)
print(hp)

pander(real_df, caption="Mann Kendall results for the Monthly GAMM residuals, ref Figure 8")
rm(Store_FwRE2)
```

# Write the monthly data as a datafile to be read in on the HPC modelling

```{r writeData}
dataOut_Month <- list()

dataOut_Month[[1]] <- flow_zoo_m 
dataOut_Month[[2]] <- rain_zoo_m 
dataOut_Month[[3]] <- maxT_zoo_m
dataOut_Month[[4]] <- flow_rain_maxT_monthly

save(dataOut_Month,file="../projectData/MonthlyDataOut.Rdata")
```